|

### 2021-03-13 supplementary_information_sciad.Rmd
Update supplementary_information_sciad.Rmd

We used `r n_scenario` sets of parameter combinations (i.e., simulation scenarios) based on the results of extensive sensitivity analysis (**Supplementary Text**). The parameter combinations comprises four landscape and eight ecological scenarios. Landscape scenarios distinguish spatial patterns of habitat heterogeneity, while ecological scenarios change ecological traits of simulated species. For landscape scenarios, we used four combinations of environmental variation at headwaters ($\sigma_h = 0.01, 1$) and the degree of local environmental noise ($\sigma_l = 0.01, 1$). The inequality of $\sigma_h$ and $\sigma_l$ controls the relationship between branching complexity and habitat heterogeneity. For ecological scenarios, we varied three ecological parameters that are relevant for dispersal and interspecific competition: dispersal distance ($\theta =$ `r unique(dat_sim$theta)`), dispersal probability ($p_d =$ `r unique(dat_sim$p_dispersal)`), and the maximum value of competition coefficients ($\alpha_{max} =$ `r unique(dat_sim$max_alpha)`). This setup results in eight ecological scenarios with different levels of dispersal and competition.

@@

We considered watersheds as independent replicates of metacommunities if the river network flows into one of the following: the ocean, a large lake/reservoir ($\ge$ 10 km² in the areal area) or a large river that may represent lentic habitats ($\ge$ 5000 km² in the watershed area). Metacommunities with $\ge$ 10 sampling sites were used in our analysis because a small sample size inflates statistical uncertainties in estimating diversity metrics (**Supplementary Text**).

@@

```{=tex}
\begin{equation}
w(v_i, v_j) = \frac{1}{1 + (v_i - v_j)^T\Omega_v^{-1}(v_i-v_j)}
\end{equation}
```
where $\Omega_v^{-1}$ is the inverse of the variance-covariance matrix of $v$. The function gives the maximum weight for a pair of $u_i$ and $u_j$ if $v_i = v_j$ (note that $v=v_i$ when $u = u_i$) while giving a less weight as the Mahalanobis distance between $v_i$ and $v_j$ increases. This property makes sense because our goal is to approximate the probability of transition from $u_i$ to $u_j$ with a common $v$ (i.e., $v_i = v_j$).

We estimated $\hat \Delta_u$ for watershed area and branching probability using the estimated coefficients $\hat \xi_k$ of the regression model explaining $\gamma$ diversity. A thousand of vectors of simulated parameters $\Theta_s$ were drawn from normal distributions with means of $\hat \xi_k$ and standard deviations of $\sigma_{\xi, k}$ (the estimated standard errors of parameters $\xi_k$). The estimated average predictive comparisons are $\hat \Delta_{u} = \frac{1}{S} \sum_{s=1}^S \hat \Delta_{u,s}$ where

>>>>

We used `r n_scenario` sets of parameter combinations (i.e., simulation scenarios) based on the results of extensive sensitivity analysis (**Supplementary Text**). The parameter combinations comprises four landscape and eight ecological scenarios. Landscape scenarios distinguish spatial patterns of habitat heterogeneity, while ecological scenarios change ecological traits of simulated species. For landscape scenarios, we used four combinations of environmental variation at headwaters ($\sigma_h = 0.01, 1$) and the degree of local environmental noise ($\sigma_l = 0.01, 1$). The inequality of $\sigma_h$ and $\sigma_l$ controls the relationship between branching complexity and habitat heterogeneity (**Figures 1A and 1B**). For ecological scenarios, we varied three ecological parameters that are relevant for dispersal and interspecific competition: dispersal distance ($\theta =$ `r unique(dat_sim$theta)`), dispersal probability ($p_d =$ `r unique(dat_sim$p_dispersal)`), and the maximum value of competition coefficients ($\alpha_{max} =$ `r unique(dat_sim$max_alpha)`). This setup results in eight ecological scenarios with different levels of dispersal and competition.

@@

We considered watersheds as independent replicates of metacommunities if the river network flows into one of the following: the ocean, a large lake/reservoir ($\ge$ 10 km² in the areal area) or a large river that may represent lentic habitats ($\ge$ 5000 km² in the watershed area). Metacommunities with $\ge$ 10 sampling sites were used in our analysis because a small sample size inflates statistical uncertainties in estimating asymptotic species richness (**Supplementary Text**).

@@

```{=tex}
\begin{equation}
w(v_i, v_j) = \frac{1}{1 + (v_i - v_j)^T\Omega_v^{-1}(v_i-v_j)}
\end{equation}
```
where $\Omega_v^{-1}$ is the inverse of the variance-covariance matrix of $v$. The function gives the maximum weight for a pair of $u_i$ and $u_j$ if $v_i = v_j$ ($v=v_i$ when $u = u_i$) while giving a less weight as the Mahalanobis distance between $v_i$ and $v_j$ increases. This property makes sense because our goal is to approximate the probability of transition from $u_i$ to $u_j$ with a common $v$ (i.e., $v_i = v_j$).

We estimated $\hat \Delta_u$ for watershed area and branching probability using the estimated coefficients $\hat \xi_k$ of the regression model explaining $\gamma$ diversity (see **Table 1**). A thousand of vectors of simulated parameters $\Theta_s$ were drawn from normal distributions with means of $\hat \xi_k$ and standard deviations of $\sigma_{\xi, k}$ (the estimated standard errors of parameters $\xi_k$). The estimated average predictive comparisons are $\hat \Delta_{u} = \frac{1}{S} \sum_{s=1}^S \hat \Delta_{u,s}$ where
